PineAPPL:\\ combining EW and QCD corrections for PDF fits

### 2020-08-17 paper/sec-conclusion.tex
added references for z+jet and first sentence of conclusions

\section{Conclusions and Outlook}
\label{sec:conclusion}

We presented a new tool called \textsc{PineAPPL}, which stores perturbative calculations from an external Monte Carlo in a PDF-independent way using interpolation grids.
This offers the advantage of fast a-posteriori convolutions with PDFs, for example to study the uncertainties coming from different PDF sets and/or the strong coupling constant $\alphas$.
Another application that we focused on is PDF fitting, where the interpolation grids constitute the theoretical input.
We tested \textsc{PineAPPL} together with \textsc{mg5\_aMC@NLO} and found a precision of \numrange{e-4}{e-5} relative to the MC result, which is more than enough for the applications described before.
Although we used \textsc{mg5\_aMC@NLO}, we note that \textsc{PineAPPL} is not tied in any way to a specific MC and can be easily interfaced with any generator.

>>>>

\section{Conclusions and Outlook}
\label{sec:conclusion}
The systematic inclusion of EW corrections in collider-physics simulations is crucial when high accuracy is required. Until
now, it was possible to include these effects only in one of the two ingredients of the physical cross section, namely
the short-distance (partonic) matrix element. For what concerns the second ingredient, PDFs, they were still extracted including
only QCD effects. In this paper, we amended to this situation, making it possible to include EW effects also in PDF fits.
We presented a new tool called \textsc{PineAPPL}, which stores perturbative calculations from an external Monte Carlo in a PDF-independent way using interpolation grids.
This offers the advantage of fast a-posteriori convolutions with PDFs, for example to study the uncertainties coming from different PDF sets and/or the strong coupling constant $\alphas$.
Another application that we focused on is PDF fitting, where the interpolation grids constitute the theoretical input.
We tested \textsc{PineAPPL} together with \textsc{mg5\_aMC@NLO} and found a precision of \numrange{e-4}{e-5} relative to the MC result, which is more than enough for the applications described before.
Although we used \textsc{mg5\_aMC@NLO}, we note that \textsc{PineAPPL} is not tied in any way to a specific MC and can be easily interfaced with any generator.

### 2020-08-17 paper/sec-results.tex
added references for z+jet and first sentence of conclusions

In this section we demonstrate the capabilities of \textsc{mg5\_aMC}+\textsc{PineAPPL} by
computing fast-interpolation grids, accurate to NLO QCD and NLO QCD+EW,
for a representative set of processes in which EW corrections may have a
sizeable effect on the accuracy of the theoretical predictions.
In order to consider some realistic kinematics for these
processes, we rely on a set of measurements that are commonly included in PDF
fits. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We describe
the settings employed for the computations, and the corresponding results for
each process.

@@

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc mg5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that involve only massless particles in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, photon-induced (PI) contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation. Monte Carlo
weights are stored as Lagrange-interpolation grids.

The central values of the renormalisation and factorisation scales, $\mu_\mathrm{R}$ and
$\mu_\mathrm{F}$, are chosen in a process-specific way, as discussed in sec.~\ref{subsec:processes_and_measurements}. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted on-the-fly in the Monte Carlo generation upon scale
variations, with the technique presented in Ref.~\cite{Frederix:2011ss}. To this purpose, we use the default \textsc{mg5\_aMC}
implementation, whereby the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed,
see equation~\eqref{eq:9pt}. However, we note that
\textsc{PineAPPL} allows the user to determine the envelope with any point
prescription, see appendix~\ref{app:pineappl-demo} for an example.

The values of the relevant physical parameters are chosen as
\begin{equation}
%\begin{aligned}
%M_\mathrm{W} &= \SI{80}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} = \SI{91.176}{\giga\electronvolt} \text{,} \quad 
m_\mathrm{t} = \SI{172.5}{\giga\electronvolt} \text{,} \quad 
%\Gamma_\mathrm{W} &= \SI{2.09}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} = \SI{2.50}{\giga\electronvolt} \text{,}
%\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{Z}$ and $m_\mathrm{t}$ are the values of the Z-boson, and
top-quark masses, respectively, and $\Gamma_\mathrm{Z}$ is the width of
the Z boson. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.

The definition of observables and cuts is process-specific, and it follows the corresponding experimental measurements,
see section~\ref{subsec:processes_and_measurements}.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematic observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination would have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
see refs.~\cite{Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order of the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of \enquote{standard} EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases {\bf MZ OTHERS???}).

@@

We focus on the following three processes: Drell--Yan lepton-pair production,
top-quark pair production, and Z-boson (lepton-pair) production with non-zero
transverse momentum in proton-proton collisions. In the following, we shall
present the experimental measurements, the process-specific settings, and
the phenomenological results for each of these processes.

When presenting the results, for each of the processes and measurements considered, we compute differential
cross sections for the observables defined in the experimental analyses in two different
ways: directly, by means of \textsc{mg5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. We
refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and EW couplings that we consider are specified for each process. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of \SI{0.1}{\permille} on the integrated cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections.

@@

We present these comparisons in
figures~\ref{fig:atlaszhighmass49fb}, \ref{fig:cmsdy2d11_bins3456}, \ref{fig:atlastop}, and \ref{fig:cmsZ13TeV}.
The format of the plots is the same across all figures. The first panel
displays the relative difference (in per mille) between the {\sc PineAPPL} and
the MC results for the central scale choice and upper/lower edges of the
scale-uncertainty envelope, for
theories accurate to both NLO QCD and NLO QCD+EW. The following three panels
present the theoretical predictions, accurate to either NLO QCD or NLO QCD+EW,
always normalised to the former; on top of the theoretical predictions, the
PDF uncertainty, the scale uncertainty and the Monte Carlo uncertainty are
displayed in turn. The relative uncertainty of the experimental data is
also shown for comparison. We shall now discuss the results for each
process and data set.

@@

\paragraph{Experimental measurements and process features.}
We select the single-differential invariant mass distribution of the
lepton pair, $M_{\ell \bar\ell}$, measured by the ATLAS experiment at a centre-of-mass
energy of \SI{7}{\tera\electronvolt} in the high-mass region
($M_{\ell\bar\ell}>\SI{116}{\giga\electronvolt}$)~\cite{Aad:2013iua}.
We also select the single-differential rapidity distribution, $y_{\ell\bar\ell}$, in slices of
the invariant mass of the lepton pair, $M_{\ell\bar\ell}$,
measured by the CMS experiment at a centre-of-mass energy of
\SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}.
These measurements are currently included as standard in the
NNPDF3.1~\cite{Ball:2017nwa} and MMHT2014~\cite{Harland-Lang:2014zoa} PDF sets,
although with appropriate kinematic cuts that remove the bins at the largest
values of invariant mass, where EW corrections become sizeable. As explained in
section~\ref{sec:pineappl-example}, 
the process has a single LO, $\mathcal{O}(\alpha^2)$; at NLO, the
QCD contribution is $\mathcal{O}(\alphas\alpha^2)$, while the EW contribution
is $\mathcal{O}(\alpha^3)$. Our NLO QCD computation includes the
$\mathcal{O}(\alpha^2)$ and $\mathcal{O}(\alphas\alpha^2)$ contributions, while
our NLO QCD+EW computation includes the $\mathcal{O}(\alpha^2)$, 
$\mathcal{O}(\alphas\alpha^2)$ and $\mathcal{O}(\alpha^3)$ contributions.
Combined QCD--EW corrections occur only at NNLO, and are therefore not
considered here. EW corrections for this process were computed in
refs.~\cite{Baur:2001ze,Dittmaier:2009cr}. The process receives contributions
from 13 (35) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.

\paragraph{Process-specific settings.}
We use a fixed value for the renormalisation and factorisation scales $\mu_\mathrm{R}=\mu_\mathrm{F}=M_\mathrm{Z}$, where $M_\mathrm{Z}$ is
the mass of the Z boson, for the ATLAS measurement, and the scale
$\mu_\mathrm{R}=\mu_\mathrm{F}=M_{\ell\bar\ell}$, where $M_{\ell\bar\ell}$ is the central value of each
invariant mass slice, for the CMS one. In the case of ATLAS, we require $p_\mathrm{T}^\ell>\SI{25}{\giga\electronvolt}$,
$|\eta_\ell|<2.5$ and $\SI{116}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$ for the transverse
momentum and the pseudorapidity of each lepton and for the invariant mass of the
lepton pair, respectively. Conversely, in the case of CMS, we require $p_\mathrm{T}^{\ell_1}>\SI{14}{\giga\electronvolt}$, $p_\mathrm{T}^{\ell_2}>\SI{9}{\giga\electronvolt}$,
$|\eta_\ell|<2.4$, $|y_{\ell\bar\ell}|<2.4$ and $\SI{20}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$
for the transverse momentum and the pseudorapidity of each lepton, and for the
rapidity and the invariant mass of the lepton pair.

\paragraph{Numerical results.}
We first consider the single-differential measurement of a lepton-pair
for large invariant masses performed by the ATLAS experiment at \SI{7}{\tera\electronvolt}.
From figure~\ref{fig:atlaszhighmass49fb} we immediately
observe that the validation of the \textsc{PineAPPL} result against the MC
result is successful. The relative difference between the two is of the order of
\SI{0.1}{\permille} at most, with negligible fluctuations across different
invariant mass bins. The agreement is similarly good irrespective of the
perturbative accuracy of the theory (NLO QCD or NLO QCD+EW) or of the scale
choice.

@@

The measurement is mainly driven by $\mathrm{q}\bar{\mathrm{q}}$ scattering: specifically, the
leading (next-to-leading) contribution comes from a $\mathrm{u}\bar{\mathrm{u}}$/$\mathrm{c}\bar{\mathrm{c}}$ ($\mathrm{d}\bar{\mathrm{d}}$/$\mathrm{s}\bar{\mathrm{s}}$)
parton luminosity, which accounts for about \SI{55}{\percent} (\SI{49}{\percent}) of the cross section
for the lowest invariant mass bins, and \SI{68}{\percent} (\SI{22}{\percent}) for the largest invariant
mass bins.\footnote{The size of these contributions may depend on the input PDF
  set. Here and in the following, we always quote results obtained from the
  \texttt{NNPDF31\_nlo\_as\_0118\_luxqed} PDF set.}
The PI contribution raises from about \SI{1.3}{\percent} in the lowest bin to
about \SI{3.6}{\percent} in the highest bin.\footnote{For the \texttt{MRST2004qed} PDF set~\cite{Martin:2004dh}, which is used to subtract PI contributions (see section~\ref{sec:doublecounting}), the value in the highest bin is roughly twice as large, \SI{6.9}{\percent} (also in absolute numbers).}
Overall, the EW corrections range between \SI{5}{\percent}
around $M_{\ell\bar\ell}\sim \SI{150}{\giga\electronvolt}$, \SIrange{2}{3}{\percent} for intermediate invariant mass
values, $\SI{150}{\giga\electronvolt}\lesssim M_{\ell\bar\ell}\lesssim \SI{700}{\giga\electronvolt}$, and
\SIrange{15}{20}{\percent} for the largest invariant mass bin, $M_{\ell\bar\ell}>\SI{1000}{\giga\electronvolt}$,
see figure~\ref{fig:atlaszhighmass49fb}. For this reason, the data points with
$M_{\ell\bar\ell}>\SI{210}{\giga\electronvolt}$ were not included in the NNPDF3.1
analysis~\cite{Ball:2017nwa}.

@@

In comparison to the PDF uncertainty, the size of the EW corrections is
always larger, especially at the boundaries of the distribution. This fact
suggests that, once included in a global fit, EW corrections will make PDFs
more accurate. In comparison to the scale
uncertainty, the size of the EW correction is similar, except for the four bins
at the largest invariant mass, where the latter is significantly larger than
the former. This fact suggests that the NNLO QCD correction is comparable
to the NLO QCD+EW correction, except at very large values of the invariant mass,
where the EW correction still dominates. This result stresses the need to
include the EW corrections in order to obtain an accurate description of the
large invariant mass bins. Finally, the Monte Carlo 
uncertainty remains negligible in comparison to the data, PDF and scale
uncertainties, and to the size of the EW correction. Our conclusions should
therefore not be affected by a generation of too few Monte Carlo events.

@@

%-------------------------------------------------------------------------------
\begin{figure}[!p]
    \centering
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin3}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin4}\\
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin5}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin6}\\
    \caption{Same as figure~\ref{fig:atlaszhighmass49fb}, but for the CMS
      double-differential Drell--Yan lepton pair measurement at a
      centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}. Displayed are
      only four of the six invariant mass bins available, respectively below the
      Z-boson mass peak, $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$, on the Z-boson mass
      peak, $\SI{60}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{120}{\giga\electronvolt}$, above the mass peak,
      $\SI{120}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{200}{\giga\electronvolt}$, and at very high invariant masses,
      $\SI{200}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$.
      Results for the slices $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ and $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ can be found in figure~\ref{fig:cmsdy2d11_bins12}.}
    \label{fig:cmsdy2d11_bins3456}
\end{figure}
%-------------------------------------------------------------------------------

>>>>

In this section we demonstrate the capabilities of \textsc{PineAPPL} by
computing fast-in\-ter\-po\-la\-tion grids, accurate to NLO QCD and NLO QCD+EW,
for a representative set of processes in which EW corrections may have a
sizeable effect on the accuracy of the theoretical predictions.
In order to consider realistic kinematics for these
processes, we rely on the measurements that are commonly included in PDF
fits. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We start by describing the
setup employed for the computation, then we discuss the features of each process, along with
the corresponding results.

@@

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that involve only massless particles in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, PI contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation. Monte Carlo
weights are stored as Lagrange-interpolation grids.

The central values of the renormalisation and factorisation scales, $\mu_\mathrm{R}$ and
$\mu_\mathrm{F}$, are chosen in a process specific way, and are discussed in sec.~\ref{subsec:processes_and_measurements}. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted on-the-fly in the Monte Carlo generation upon scale
variations, with the technique presented in Ref.~\cite{Frederix:2011ss}. To this purpose, we use the default \textsc{MG5\_aMC}
implementation, whereby the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed. However, we note that
\textsc{PineAPPL} allows the user to determine the envelope with any point
prescription, see appendix~\ref{app:pineappl-demo} for an example.

The values of the relevant physical parameters are chosen as
\begin{equation}
\begin{aligned}
M_\mathrm{W} &= \SI{80}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} &= \SI{91.176}{\giga\electronvolt} \text{,} \quad &
m_\mathrm{t} &= \SI{172.5}{\giga\electronvolt} \text{,} \quad \\
\Gamma_\mathrm{W} &= \SI{2.09}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} &= \SI{2.50}{\giga\electronvolt} \text{,}
\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{W}$, $M_\mathrm{Z}$ and $m_\mathrm{t}$ are the values of the W-boson, Z-boson, and
top quark masses, respectively, and $\Gamma_\mathrm{W}$ and $\Gamma_\mathrm{Z}$ are the width of
the W boson and of the Z boson, respectively. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.
The definition of observables and cuts is process-specific, and it follows the corresponding experimental measurements.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematics observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination would have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
see refs.~\cite{Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order as the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of ``standard'' EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases {\bf MZ OTHERS???}).

@@

We focus on the following three processes: Drell--Yan lepton-pair production,
top-quark pair production, and Z-boson (lepton-pair) production with non-zero
transverse momentum in proton-proton collisions. For each of these processes,
we present the experimental measurements, process-specific settings, and phenomenological results.

When presenting the results, for each of the processes and measurements considered, we compute the expectation
value of the corresponding observable for each kinematic bin in two different
ways: directly, by means of \textsc{MG5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. We
will refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and electroweak couplings that we consider are specified for each process. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of a fraction of per mille on the total cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections.

@@

We present these comparisons in
figures~\ref{fig:atlaszhighmass49fb}--\ref{fig:cmsZ13TeV}.
The format of the plots is the same across all figures. The first panel
displays the relative difference (in per mille) between the {\sc PineAPPL} and
the MC results for the central scale choice and upper/lower edges of the
scale-uncertainty envelope, for
theories accurate to both NLO QCD and NLO QCD+EW. The following three panels
present the theoretical predictions, accurate to either NLO QCD or NLO QCD+EW,
always normalised to the former; on top of the theoretical predictions, the
PDF uncertainty, the scale uncertainty and the Monte Carlo uncertainty are
displayed in turn. The relative uncertainty of the experimental data is
also shown for comparison. We shall now discuss the results for each
process and data set.

@@

\paragraph{Experimental measurements and process features.}
We select the distribution, single-differential in the invariant mass of the
lepton pair, $M_{\ell \bar\ell}$, measured by the ATLAS experiment at a centre-of-mass
energy of \SI{7}{\tera\electronvolt} in the high-mass region
($M_{\ell\bar\ell}>\SI{116}{\giga\electronvolt}$)~\cite{Aad:2013iua}.
We also select the distributions, double-differential in the rapidity and in
the invariant mass of the lepton pair, $y_{\ell\bar\ell}$ and $M_{\ell\bar\ell}$,
measured by the CMS experiment at a centre-of-mass energy of
\SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}.
These measurements are currently included as standard in the
NNPDF3.1~\cite{Ball:2017nwa} and MMHT2014~\cite{Harland-Lang:2014zoa} PDF sets,
although with appropriate kinematic cuts that remove the bins at the largest
values of invariant mass, where EW corrections become sizeable. As explained in
section~\ref{sec:pineappl-example}, 
the process has a single LO, $\mathcal{O}(\alpha^2)$; at NLO, the
QCD contribution is $\mathcal{O}(\alphas\alpha^2)$, while the EW contribution
is $\mathcal{O}(\alpha^3)$. Our NLO QCD computation includes the
$\mathcal{O}(\alpha^2)$ and $\mathcal{O}(\alphas\alpha^2)$ contributions, while
our NLO QCD+EW computation includes the $\mathcal{O}(\alpha^2)$, 
$\mathcal{O}(\alphas\alpha^2)$ and $\mathcal{O}(\alpha^3)$ contributions.
Mixed-coupling QCD-EW corrections occur only at NNLO, and are therefore not
considered here. EW corrections for this process were computed in
refs.~\cite{Baur:2001ze,Dittmaier:2009cr}. The process receives contributions
from 13 (35) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.

\paragraph{Process-specific settings.}
We use a fixed value for the renormalisation and factorisation scales $\mu_\mathrm{R}=\mu_\mathrm{F}=M_\mathrm{Z}$, where $M_\mathrm{Z}$ is
the mass of the Z boson, for the ATLAS measurement, and the scale
$\mu_\mathrm{R}=\mu_\mathrm{F}=M_{\ell\bar\ell}$, where $M_{\ell\bar\ell}$ is the central value of each
invariant mass bin, for the CMS one. In the case of ATLAS, we require $p_\mathrm{T}^\ell>\SI{25}{\giga\electronvolt}$,
$|\eta_\ell|<2.5$ and $\SI{116}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$ for the transverse
momentum and the pseudorapidity of each lepton and for the invariant mass of the
lepton pair, respectively. Conversely, in the case of CMS, we require $p_\mathrm{T}^{\ell_1}>\SI{14}{\giga\electronvolt}$, $p_\mathrm{T}^{\ell_2}>\SI{9}{\giga\electronvolt}$,
$|\eta_\ell|<2.4$, $|\eta_{\ell\bar\ell}|<2.4$ and $\SI{20}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$
for the transverse momentum and the pseudorapidity of each lepton, and for the
pseudorapidity and the invariant mass of the lepton pair.

\paragraph{Numerical results.}
We first consider the single-differential measurement at a high lepton-pair
invariant mass performed by the ATLAS experiment at \SI{7}{\tera\electronvolt}.
From figure~\ref{fig:atlaszhighmass49fb} we immediately
observe that the validation of the \textsc{PineAPPL} result against the MC
result is successful. The relative difference between the two is of the order
\SI{0.1}{\permille} at most, with negligible fluctuations across different
invariant mass bins. The agreement is similarly good irrespective of the
perturbative accuracy of the theory (NLO QCD or NLO QCD+EW) or of the scale
choice.

@@

The measurement is mainly driven by $\mathrm{q}\bar{\mathrm{q}}$ scattering: specifically, the
leading (next-to-leading) contribution comes from a $\mathrm{u}\bar{\mathrm{u}}$ ($\mathrm{d}\bar{\mathrm{d}}$)
parton luminosity, which accounts for about \SI{55}{\percent} (\SI{49}{\percent}) of the cross section
for the lowest invariant mass bins, and \SI{68}{\percent} (\SI{22}{\percent}) for the largest invariant
mass bins.\footnote{The size of these contributions may depend on the input PDF
  set. Here and in the following, we always quote results obtained from the
  \texttt{NNPDF31\_nlo\_as\_0118\_luxqed} PDF set.}
The PI contribution raises from about \SI{1.3}{\percent} in the lowest bin to
about \SI{3.6}{\percent} in the highest bin. Overall, the EW corrections range between \SI{5}{\percent}
around $M_{\ell\bar\ell}\sim \SI{150}{\giga\electronvolt}$, \SIrange{2}{3}{\percent} for intermediate invariant mass
values, $\SI{150}{\giga\electronvolt}\lesssim M_{\ell\bar\ell}\lesssim \SI{700}{\giga\electronvolt}$, and
\SIrange{15}{20}{\percent} for the largest invariant mass bin, $M_{\ell\bar\ell}>\SI{1000}{\giga\electronvolt}$,
see figure~\ref{fig:atlaszhighmass49fb}. For this reason, the data points with
$M_{\ell\bar\ell}>\SI{210}{\giga\electronvolt}$ were not included in the NNPDF3.1
analysis~\cite{Ball:2017nwa}.

@@

In comparison to the PDF uncertainty, the size of the EW corrections is
always larger, especially at the boundaries of the distribution. This fact
suggests that, once included in a global fit, EW corrections could recast the
relative weight of each data set included in the fit, and possibly lead to
an improvement in their overall description. In comparison to the scale
uncertainty, the size of the EW correction is similar, except for the four bins
at the largest invariant mass, where the latter is significantly larger than
the former. The size of the scale variation is the same for the QCD and the
QCD+EW theories. These facts suggest that the NNLO QCD correction is comparable
to the NLO QCD+EW correction, except at very large values of the invariant mass,
where the EW correction still dominates. This result stresses the need to
include the EW corrections in order to obtain an accurate description of the
large invariant mass bins. Finally, the Monte Carlo statistical
uncertainty remains negligible in comparison to the data, PDF and scale
uncertainties, and to the size of the EW correction. Our conclusions should
therefore not be affected by a too low number of Monte Carlo events generated.

@@

%-------------------------------------------------------------------------------
\begin{figure}[!p]
    \centering
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin3}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin4}\\
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin5}%
    \includegraphics[width=0.46\textwidth]{figures/pineappl_CMSDY2D11_bin6}\\
    \caption{Same as figure~\ref{fig:atlaszhighmass49fb}, but for the CMS
      double-differential Drell--Yan lepton pair measurement at a
      centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}. Displayed are
      only two of the six invariant mass bins available, respectively below the
      Z-boson mass peak, $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$, on the Z-boson mass
      peak, $\SI{60}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{120}{\giga\electronvolt}$, above the mass peak,
      $\SI{120}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{200}{\giga\electronvolt}$, and at very high invariant masses,
      $\SI{200}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$.
      Results for the slices $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ and $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ can be found in figure~\ref{fig:cmsdy2d11_bins12}.}
    \label{fig:cmsdy2d11_bins3456}
\end{figure}
%-------------------------------------------------------------------------------

As in the case of the ATLAS high-mass Drell--Yan measurement, the CMS
measurement is also dominated by $\mathrm{q}\bar{\mathrm{q}}$ scattering. The leading
(next-to-leading) contribution to the $\SI{45}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{60}{\giga\electronvolt}$ invariant
mass bin comes from
the $\mathrm{u}\bar{\mathrm{u}}$ ($\mathrm{d}\bar{\mathrm{d}}$) parton luminosity, which accounts for about \SI{70}{\percent}
(\SI{22}{\percent}) of the double differential cross section, with small fluctuations across
the rapidity range. The PI contribution decreases from about \SI{4}{\percent} at zero
rapidity to \SI{1.5}{\percent} in the largest rapidity bin. The situation is slightly
different in the $\SI{60}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{120}{\giga\electronvolt}$ invariant mass bin, where the
leading (next-to-leading) contribution comes instead from the $\mathrm{d}\bar{\mathrm{d}}$
($\mathrm{u}\bar{\mathrm{u}}$) parton luminosity, which accounts for about \SI{60}{\percent} (\SI{44}{\percent}) of the
double differential cross section at small rapidities, and for about \SI{56}{\percent} (\SI{50}{\percent})
at large rapidities. In the remaining two invariant mass bins, the leading
(next-to-leading) contribution comes again from the $\mathrm{u}\bar{\mathrm{u}}$ ($\mathrm{d}\bar{\mathrm{d}}$)
parton luminosity, which accounts for about \SIrange{69}{95}{\percent} (\SIrange{38}{30}{\percent}) and
\SIrange{57}{70}{\percent} (\SIrange{48}{34}{\percent}) of the cross section, respectively for
$\SI{120}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{200}{\giga\electronvolt}$ and $\SI{200}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$ in the
corresponding rapidity intervals; PI contributions range between \SIrange{7.3}{1.2}{\percent}
and \SIrange{3.7}{0.6}{\percent} in the two invariant mass bins, respectively, for increasing
rapidity.
